---
title: "Working model without nematodes, with (literature parameters, PRETTY interactive UI, calibrated parameters, literature state variables, 13C version)."
author: "Lucy"
date: "28/04/2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model setup

Loading the libraries necessary, deSolve for the run of the model and shiny for the interactive UI.

```{r}
##------------------------------------------------block library loading-----------------------------------------------------
require(base)
require(datasets)
require(deSolve)
require(graphics)
require(grDevices)
require(methods)
require(shiny)
require(shinyWidgets)
require(stats)
require(utils)
##---------------------------------------------end of block library loading-------------------------------------------------
```

Definition of initial state variables, initial parameters for the default model run.

```{r}
##----------------------------------------block state variables and parameters----------------------------------------------

#----------------------------------------initial natural abundance state variables------------------------------------------
#C.INI <- 15 (300/180*6 or 12-14) - this is an old value which is equal to unlabeled compost
#previously looked like this: state <- c(BACTERIA = 0.022*C.INI, FUNGI = 0.0177*C.INI, A.BISPORUS = 0.01*C.INI, SUGARS = 0.26*C.INI, COMPOST = C.INI, CO2 = 0)
#state variables were made to be a list, not a vector using c(), they are unlisted for the ode function where necessary as ode only takes y which is numeric, but UI takes only state variables as a named list
#for the same reason, values of state variables are now numbers instead of percentages, but the values are based on COMPOST = 15 still, just recalculated to fit right from previous percentages

#excess 13C as calculated experimentally from PLFAs and other experiments, can be seen as amount of substance fractions x13C = COMPARTMENT.13C/COMPARTMENT
#in mmol/g compost dry weight
e.bac      = 0.001042
e.fun      = 8.1e-5
e.bisp     = 8.1e-5 #2.88e-9
e.sugars   = 0.011421
e.co2      = 2.31e-7


state     <- list(BACTERIA       = 1.475,
                  BACTERIA.13C   = 1.475*e.bac,
                  FUNGI          = 1.6, 
                  FUNGI.13C      = 1.6*e.fun,
                  A.BISPORUS     = 0.008357, 
                  A.BISPORUS.13C = 0.008357*e.bisp,
                  SUGARS         = 8.58222, 
                  SUGARS.13C     = 8.58222*e.sugars,
                  COMPOST        = 25, 
                  CO2            = 0.00001,
                  CO2.13C        = 0.00001*e.co2
)

#-------------------------------------parameters that comprise rate laws of processes go here--------------------------------

parms <- list(k1.deg.bac          = 0.05,             #0.05 - 0.15, literature value
              k2.deg.fun          = 0.05,             #0.05 - 0.15, literature value
              k3.deg.bisp         = 0.01,             #0.075 - 0.13, literature value
              k4.bac.uptake       = 0.1,              #0.04 - 0.1, literature value
              k5.fun.uptake       = 0.05,             #0.01 - 0.05, literature value
              k6.bisp.uptake      = 0.15,             #needs calibration, actually unknown, placeholder value used
              k7.fun.killing.bac  = 0.2,              #needs calibration, actually unknown, placeholder value used
              k8.bisp.killing.bac = 0.1,              #needs calibration, actually unknown, placeholder value used
              k9.bisp.killing.fun = 0.5,              #needs calibration, actually unknown, placeholder value used
              k10.bac.mort        = 0.24,             #0.24 - 0.72, literature value
              k11.fun.mort        = 0.1,              #0.01 - 0.1, literature value
              k12.bisp.mort       = 0.01,             #literature value
              kBISPORUS           = 0.5,              #needs calibration, actually unknown, placeholder value used
              MAX.A.BISPORUS      = 1,                #needs calibration, actually unknown, placeholder value used
              kSUGARS.bac         = 0.4,              #needs calibration, actually unknown, placeholder value used
              kSUGARS.fun         = 0.4,              #needs calibration, actually unknown, placeholder value used
              kSUGARS.bisp        = 0.2,              #needs calibration, actually unknown, placeholder value used
              Ef.fun              = 0.3,              #literature value
              Ef.bac              = 0.3,              #literature value
              Ef.bisp             = 0.4               #literature value
)

##---------------------------------------end of block state variables, parameters---------------------------------------------
```

Definition of the model function and the output of the model.

```{r}
##-------------------------------------------block model definition-----------------------------------------------------------
WHITEBUTTON <- function(t, state, parameters) { 
  with(as.list(c(state, parameters)),{
    
      x13C.bac            <- BACTERIA.13C/BACTERIA
      x13C.fun            <- FUNGI.13C/FUNGI
      x13C.bisp           <- A.BISPORUS.13C/A.BISPORUS
      x13C.sugars         <- SUGARS.13C/SUGARS
      x13C.co2            <- CO2.13C/CO2
 
#------------------------Rate expressions - all in units of [mmolC/m3/day] or [mmolC/g compost] go here-----------------------
    
      R1.deg.bac          <- k1.deg.bac * BACTERIA
      R2.deg.fun          <- k2.deg.fun * FUNGI
      R3.deg.bisp         <- k3.deg.bisp * A.BISPORUS
      R4.bac.uptake       <- k4.bac.uptake * SUGARS/(SUGARS+kSUGARS.bac) * BACTERIA * kBISPORUS/(A.BISPORUS+kBISPORUS)
      R5.fun.uptake       <- k5.fun.uptake * SUGARS/(SUGARS+kSUGARS.fun) * FUNGI * kBISPORUS/(A.BISPORUS+kBISPORUS)
      R6.bisp.uptake      <- k6.bisp.uptake * SUGARS/(SUGARS+kSUGARS.bisp) * A.BISPORUS * (1-A.BISPORUS/MAX.A.BISPORUS)
      R7.fun.killing.bac  <- k7.fun.killing.bac * BACTERIA * FUNGI
      R8.bisp.killing.bac <- k8.bisp.killing.bac * BACTERIA * A.BISPORUS
      R9.bisp.killing.fun <- k9.bisp.killing.fun * FUNGI * A.BISPORUS
      R10.bac.mort        <- k10.bac.mort * BACTERIA
      R11.fun.mort        <- k11.fun.mort * FUNGI
      R12.bisp.mort       <- k12.bisp.mort * A.BISPORUS
      
#--------------------------------Mass balances [molC/m3/day or molC/kg/day] go here-------------------------------------------
      
#mass balances are first given for unlabeled then for labelled model versions, rates are multiplied w/ fraction of the source
      
      dBACTERIA       <- Ef.bac*R4.bac.uptake - 
                         R7.fun.killing.bac -
                         R8.bisp.killing.bac - 
                         R10.bac.mort 
                     
      dBACTERIA.13C   <- Ef.bac*R4.bac.uptake*x13C.sugars - 
                         R7.fun.killing.bac*x13C.bac -
                         R8.bisp.killing.bac*x13C.bac - 
                         R10.bac.mort*x13C.bac
      
      dFUNGI          <- Ef.fun*R5.fun.uptake - 
                         R9.bisp.killing.fun - 
                         R11.fun.mort
      
      dFUNGI.13C      <- Ef.fun*R5.fun.uptake*x13C.sugars - 
                         R9.bisp.killing.fun*x13C.fun - 
                         R11.fun.mort*x13C.fun
      
      dA.BISPORUS     <- Ef.bisp*R6.bisp.uptake - 
                         R12.bisp.mort
      
      dA.BISPORUS.13C <- Ef.bisp*R6.bisp.uptake*x13C.sugars -
                         R12.bisp.mort*x13C.bisp
      
      dSUGARS         <- R1.deg.bac + 
                         R2.deg.fun + 
                         R3.deg.bisp -
                         R4.bac.uptake - 
                         R5.fun.uptake - 
                         R6.bisp.uptake + 
                         R7.fun.killing.bac + 
                         R8.bisp.killing.bac + 
                         R9.bisp.killing.fun +
                         R10.bac.mort + 
                         R11.fun.mort + 
                         R12.bisp.mort
      
      dSUGARS.13C     <- R1.deg.bac + 
                         R2.deg.fun + 
                         R3.deg.bisp -
                         R4.bac.uptake*x13C.sugars - 
                         R5.fun.uptake*x13C.sugars - 
                         R6.bisp.uptake*x13C.sugars + 
                         R7.fun.killing.bac*x13C.bac + 
                         R8.bisp.killing.bac*x13C.bac + 
                         R9.bisp.killing.fun*x13C.fun +
                         R10.bac.mort*x13C.bac + 
                         R11.fun.mort*x13C.fun + 
                         R12.bisp.mort*x13C.bisp
      
      #no label in compost!
      dCOMPOST        <- -R1.deg.bac - R2.deg.fun - R3.deg.bisp
        
      dCO2            <- (1-Ef.bac)*R4.bac.uptake + 
                         (1-Ef.fun)*R5.fun.uptake + 
                         (1-Ef.bisp)*R6.bisp.uptake

      dCO2.13C        <- (1-Ef.bac)*R4.bac.uptake*x13C.sugars + 
                         (1-Ef.fun)*R5.fun.uptake*x13C.sugars + 
                         (1-Ef.bisp)*R6.bisp.uptake*x13C.sugars
      
      TOTAL_C         <- COMPOST + BACTERIA     + SUGARS     + FUNGI     + A.BISPORUS     + CO2
      TOTAL_C.13C     <-           BACTERIA.13C + SUGARS.13C + FUNGI.13C + A.BISPORUS.13C + CO2.13C
      
#----------------------------------------------Model output--------------------------------------------------------------------
      
      return (list(c(dBACTERIA, dBACTERIA.13C, dFUNGI, dFUNGI.13C, dA.BISPORUS, dA.BISPORUS.13C, dSUGARS, dSUGARS.13C, dCOMPOST, dCO2, dCO2.13C),
                   TOTAL_C         = TOTAL_C, 
                   TOTAL_C.13C     = TOTAL_C.13C
                   )
              )
      })
}

##-------------------------------------------end of block model definition------------------------------------------------------
```

Model is ran and the results are plotted separately - argument which defines the output to model change of individually selected parameters that wish to be tracked through the duration of the dynamic model.

```{r}
##---------------------------------------------block model run with deSolve-----------------------------------------------------

outtimes     <- seq(from = 0, to = 40, length.out = 1000)
model_output <- ode(y = unlist(state), parms = parms, func = WHITEBUTTON, times = outtimes) #unlisted to use with ode, as mentioned

##-------------------------------------------end of block model run with deSolve------------------------------------------------
```

```{r, fig.height=4, fig.width=8}
##---------------------------------------------block model output plotting------------------------------------------------------

plot(model_output, xlab="time (days)", mfrow=c(1,3), which=c("BACTERIA", "BACTERIA.13C", "FUNGI", "FUNGI.13C", "A.BISPORUS", "A.BISPORUS.13C"))
plot(model_output, xlab="time (days)", mfrow=c(1,3), which=c("SUGARS", "SUGARS.13C", "COMPOST", "CO2", "CO2.13C"))
plot(model_output, xlab="time (days)", mfrow=c(1,3), which=c("TOTAL_C", "TOTAL_C.13C"))

##---------------------------------------------block model output plotting------------------------------------------------------
```

Lastly, some code can be added in order to make the output interactive and open im a browser, where parameters can be edited in real time.

```{r}
##---------------------------------------------block interactive user interface-------------------------------------------------

#-------------------------------------------------------------------------------------------------------------------------------

state       <- state
times       <- outtimes
defaultprog <- ode(y = unlist(state), times = outtimes, func = WHITEBUTTON, parms = parms) #unlisted to use with ode

#------------------------------------------------ui-side part of the code-------------------------------------------------------

ui <- fluidPage(fluid = TRUE,

#title, centered

  titlePanel( 
    h1("Compost Food Web Interactive Simulation", windowTitle = "Compost Food Web Interactive Simulation", align = "center")
  ),
  
#UI output
  uiOutput("WB"),

#sidebar and tabs
  sidebarLayout(

#aligns the plot at the center and resizes it to a nice size for both browser and RStudio launcher    
   mainPanel(
     fluidRow(
          column(12, offset = 2, align="center",
      plotOutput("PlotWB", height = "700px", width = "700px")
          )
     )
   ),
      
 #creates UI using javascript conditionalPanel tests, bunches terms together under a single tab based on conditions
   CreateUI <- renderUI({  
    sidebarPanel(width = 12,
       tabsetPanel(id = "tabselected",
                   tabPanel(strong("degradation"),        value = 1, helpText("degradation of polysaccharides from compost")),
                   tabPanel(strong("uptake"),             value = 2, helpText("mono-sugar uptake")),
                   tabPanel(strong("predation"),          value = 3, helpText("predation rates")),
                   tabPanel(strong("mortality"),          value = 4, helpText("mortality rates")),
                   tabPanel(strong("Monod"),              value = 5, helpText("Monod (Michaelis-Menten) kinetics")),
                   tabPanel(strong("efficiency"),         value = 6, helpText("uptake efficiency of sugars")),
                   tabPanel(strong("initial states, L-"), value = 7, helpText("initial state variables (amounts and biomasses), natural abundance")),
                   tabPanel(strong("initial states, L+"), value = 8, helpText("labelled compartments - smaller number = smaller excess, larger number = larger excess than default"))
       ),

#be VERY CAREFUL with the indentation of conditional panels, otherwise nothing will ever work          
       conditionalPanel(condition = "input.tabselected == 1",
            fluidRow(
                  column(12,
                         sliderInput(inputId="k1.deg.bac", label = "degradation into sugars by bacteria (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k1.deg.bac),
                         sliderInput(inputId="k2.deg.fun", label = "degradation into sugars by fungi (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k2.deg.fun),
                         sliderInput(inputId="k3.deg.bisp",label = "degradation into sugars by A. bisporus (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k3.deg.bisp)       
                  ),
            ),
       ),

       conditionalPanel(condition = "input.tabselected == 2",
             fluidRow(
                   column(12,
                         sliderInput(inputId="k4.bac.uptake", label = "maximum carb uptake rate by bacteria (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k4.bac.uptake),
                         sliderInput(inputId="k5.fun.uptake", label = "maximum carb uptake rate by fungi (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k5.fun.uptake),
                         sliderInput(inputId="k6.bisp.uptake", label = "maximum carb uptake rate by A. bisporus (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k6.bisp.uptake)     
                   ),
             ),
       ),

       conditionalPanel(condition = "input.tabselected == 3",
             fluidRow(
                   column(12,
                         sliderInput(inputId="k7.fun.killing.bac", label = "maximum predation rate constant, fungi > bacteria (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k7.fun.killing.bac),
                         sliderInput(inputId="k8.bisp.killing.bac", label = "maximum predation rate constant, A. bisporus > bacteria (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k8.bisp.killing.bac),
                         sliderInput(inputId="k9.bisp.killing.fun", label = "maximum predation rate constant, A. bisporus > fungi (/d)", min = 0.01, max = 1.0, step = 0.01, value = parms$k9.bisp.killing.fun)   
                   ),
             ),
       ),
            
       conditionalPanel(condition = "input.tabselected == 4",
             fluidRow(
                   column(12,
                         sliderInput(inputId="k10.bac.mort", label = "bacterial linear mortality constant [(mmol C per cubic meter)/d]", min = 0.01, max = 1.0, step = 0.01, value = parms$k10.bac.mort),
                         sliderInput(inputId="k11.fun.mort", label = "fungal linear mortality constant [(mmol C per cubic meter)/d]",   min = 0.01, max = 1.0, step = 0.01, value = parms$k11.fun.mort),
                         sliderInput(inputId="k12.bisp.mort", label = "A. bisporus linear mortality constant [(mmol C per cubic meter)/d]", min = 0.01, max = 1.0, step = 0.01, value = parms$k12.bisp.mort) 
                   ),
             ),
       ),        

       conditionalPanel(condition = "input.tabselected == 5",
             fluidRow(
                   column(12,
                         sliderInput(inputId="kBISPORUS", label = "Monod constant for A. bisporus (mmol C per cubic meter)", min = 0.01, max = 1.0, step = 0.01, value = parms$kBISPORUS),
                         sliderInput(inputId="MAX.A.BISPORUS", label = "maximum unlimited A. bisporus growth (mmol C per cubic meter)", min = 0.01, max = 5.0, step = 0.1, value = parms$MAX.A.BISPORUS),
                         sliderInput(inputId="kSUGARS.bac", label = "bacterial growth limit (mmol C per cubic meter)", min = 0.01, max = 2.0, step = 0.01, value = parms$kSUGARS.bac),
                         sliderInput(inputId="kSUGARS.fun", label = "fungal growth limit (mmol C per cubic meter)", min = 0.01, max = 2.0, step = 0.01, value = parms$kSUGARS.fun),
                         sliderInput(inputId="kSUGARS.bisp", label = "A. bisporus growth limit (mmol C per cubic meter)", min = 0.01, max = 2.0, step = 0.01, value = parms$kSUGARS.bisp) 
                   ),
             ),
       ),

       conditionalPanel(condition = "input.tabselected == 6",
             fluidRow(
                   column(12,
                         sliderInput(inputId="Ef.bac", label = "%C assimilated by bacteria due to eating (-)", min = 0.01, max = 1.0, step = 0.01, value = parms$Ef.bac),
                         sliderInput(inputId="Ef.fun", label = "%C assimilated by fungi due to eating (-)", min = 0.01, max = 1.0, step = 0.01, value = parms$Ef.fun),
                         sliderInput(inputId="Ef.bisp", label = "%C assimilated by A. bisporus due to eating (-)", min = 0.01, max = 1.0, step = 0.01, value = parms$Ef.bisp)
                   ),
             ),
       ),
       
       conditionalPanel(condition = "input.tabselected == 7",
             fluidRow(
                   column(12,
                         sliderInput(inputId="BACTERIA", label = "starting bacterial biomass", min = 0.1, max = 3, step = 0.001, value = state$BACTERIA),                                                                                        #1.475 is the default value
                         sliderInput(inputId="FUNGI", label = "starting fungal biomass", min = 0.1, max = 3, step = 0.01, value = state$FUNGI),                                                                                                  #1.6 is the default value
                         sliderInput(inputId="A.BISPORUS", label = "starting A. bisporus biomass", min = 0.001, max = 0.01, step = 0.001, value = state$A.BISPORUS),                                                                             #0.008357 is the default value
                         sliderInput(inputId="SUGARS", label = "starting amount of sugars in compost", min = 3, max = 13, step = 0.1, value = state$SUGARS),                                                                                     #8.58222 is the default value
                         sliderInput(inputId="COMPOST", label = "starting amount of polysaccharides in compost", min = 10, max = 40, step = 1, value = state$COMPOST),                                                                           #25 is the default value
                         sliderInput(inputId="CO2", label = "respired carbon dioxide due to biological activities", min = 0, max = 5, step = 0.01, value = state$CO2)                                                                            #0 is the default value
                   ),
             ),
       ),

       conditionalPanel(condition = "input.tabselected == 8",
             fluidRow(
                   column(12,
                         numericInput(inputId="BACTERIA.13C", label = "labelled bacterial biomass", value = 0.00153695, min = 0.0001, max = 1, step = 0.0001),                                                                                  #0.00153695 is the default value
                         numericInput(inputId="FUNGI.13C", label = "labelled fungal biomass", value = 0.0001296, min = 0.00001, max = 1, step = 0.00001),                                                                                       #0.0001296 is the default value
                         numericInput(inputId="A.BISPORUS.13C", label = "labelled A. bisporus biomass", value = 0.0001296, min = 0.00001, max = 1, step = 0.00001),                                                                            #2.406816e-11 is the real value, but we try with mushrooms value here
                         numericInput(inputId="SUGARS.13C", label = "starting amount of labelled sugars in compost", value = 0.09801753, min = 0.001, max = 1, step = 0.001),                                                               #0.09801753 is the default value
                         numericInput(inputId="CO2.13C", label = "respired labelled carbon dioxide due to biological activities", value = 2.31e-12, min = 1e-12, max = 1e-13, step = 1e-13)                                                   #2.31e-12 is the default value
                   ),
             ),
       ),

   actionButton (inputId="resetButton", label="Reset Parameters"),
    
   checkboxInput(inputId="defaultRun",  label=strong("Add default run?"), value=TRUE),
   br(),
   )

    })
  )
)

#-----------------------------------------------server-side part of the code----------------------------------------------------

server <- function(input, output, session) {

#ties the UI coded above to the server  
  output$WB <- renderUI({  
    CreateUI})

#------------------------------------------------reset button definition--------------------------------------------------------
  
  observeEvent(input$resetButton, {
    #resets parameters
    updateNumericInput(session, "k1.deg.bac",             value = parms$k1.deg.bac)
    updateNumericInput(session, "k2.deg.fun",             value = parms$k2.deg.fun)
    updateNumericInput(session, "k3.deg.bisp",            value = parms$k3.deg.bisp)
    updateNumericInput(session, "k4.bac.uptake",          value = parms$k4.bac.uptake)
    updateNumericInput(session, "k5.fun.uptake",          value = parms$k5.fun.uptake)
    updateNumericInput(session, "k6.bisp.uptake",         value = parms$k6.bisp.uptake)
    updateNumericInput(session, "k7.fun.killing.bac",     value = parms$k7.fun.killing.bac)
    updateNumericInput(session, "k8.bisp.killing.bac",    value = parms$k8.bisp.killing.bac)
    updateNumericInput(session, "k9.bisp.killing.fun",    value = parms$k9.bisp.killing.fun)
    updateNumericInput(session, "k10.bac.mort",           value = parms$k10.bac.mort)
    updateNumericInput(session, "k11.fun.mort",           value = parms$k11.fun.mort)
    updateNumericInput(session, "k12.bisp.mort",          value = parms$k12.bisp.mort)
    updateNumericInput(session, "kBISPORUS",              value = parms$kBISPORUS)
    updateNumericInput(session, "MAX.A.BISPORUS",         value = parms$MAX.A.BISPORUS)
    updateNumericInput(session, "kSUGARS.bac",            value = parms$kSUGARS.bac)
    updateNumericInput(session, "kSUGARS.fun",            value = parms$kSUGARS.fun)
    updateNumericInput(session, "kSUGARS.bisp",           value = parms$kSUGARS.bisp)
    updateNumericInput(session, "Ef.bac",                 value = parms$Ef.bac)
    updateNumericInput(session, "Ef.fun",                 value = parms$Ef.fun)
    updateNumericInput(session, "Ef.bisp",                value = parms$Ef.bisp)
    #resets state variables
    updateNumericInput(session, "BACTERIA",               value = state$BACTERIA)
    updateNumericInput(session, "BACTERIA.13C",           value = state$BACTERIA.13C)
    updateNumericInput(session, "FUNGI",                  value = state$FUNGI)
    updateNumericInput(session, "FUNGI.13C",              value = state$FUNGI.13C)
    updateNumericInput(session, "A.BISPORUS",             value = state$A.BISPORUS)
    updateNumericInput(session, "A.BISPORUS.13C",         value = state$A.BISPORUS.13C)
    updateNumericInput(session, "SUGARS",                 value = state$SUGARS)
    updateNumericInput(session, "SUGARS.13C",             value = state$SUGARS.13C)
    updateNumericInput(session, "COMPOST",                value = state$COMPOST)
    updateNumericInput(session, "CO2",                    value = state$CO2)
    updateNumericInput(session, "CO2.13C",                value = state$CO2.13C)
  })

#-------------------acquires model parameters and state variables which are defined in the user interface-----------------------
  
  getparms <- reactive( {
    newparms                        <- parms 
    newparms$k1.deg.bac             <- input$k1.deg.bac
    newparms$k2.deg.fun             <- input$k2.deg.fun   
    newparms$k3.deg.bisp            <- input$k3.deg.bisp 
    newparms$k4.bac.uptake          <- input$k4.bac.uptake  
    newparms$k5.fun.uptake          <- input$k5.fun.uptake
    newparms$k6.bisp.uptake         <- input$k6.bisp.uptake 
    newparms$k7.fun.killing.bac     <- input$k7.fun.killing.bac
    newparms$k8.bisp.killing.bac    <- input$k8.bisp.killing.bac 
    newparms$k9.bisp.killing.fun    <- input$k9.bisp.killing.fun 
    newparms$k10.bac.mort           <- input$k10.bac.mort
    newparms$k11.fun.mort           <- input$k11.fun.mort
    newparms$k12.bisp.mort          <- input$k12.bisp.mort 
    newparms$kBISPORUS              <- input$kBISPORUS
    newparms$MAX.A.BISPORUS         <- input$MAX.A.BISPORUS 
    newparms$kSUGARS.bac            <- input$kSUGARS.bac
    newparms$kSUGARS.fun            <- input$kSUGARS.fun 
    newparms$kSUGARS.bisp           <- input$kSUGARS.bisp
    newparms$Ef.bac                 <- input$Ef.bac 
    newparms$Ef.fun                 <- input$Ef.fun
    newparms$Ef.bisp                <- input$Ef.bisp 
    newparms
  })

  getstate <- reactive( {
    newstate                        <- state
    newstate$BACTERIA               <- input$BACTERIA
    newstate$BACTERIA.13C           <- input$BACTERIA.13C
    newstate$FUNGI                  <- input$FUNGI 
    newstate$FUNGI.13C              <- input$FUNGI.13C 
    newstate$A.BISPORUS             <- input$A.BISPORUS
    newstate$A.BISPORUS.13C         <- input$A.BISPORUS.13C
    newstate$SUGARS                 <- input$SUGARS
    newstate$SUGARS.13C             <- input$SUGARS.13C
    newstate$COMPOST                <- input$COMPOST 
    newstate$CO2                    <- input$CO2
    newstate$CO2.13C                <- input$CO2.13C
    newstate
  })
  
#-------------------------------------------------definition of the plot output-------------------------------------------------
  
  output$PlotWB <- renderPlot({

   newparms     <- getparms()
   newstate     <- getstate()
   out          <- ode(y = unlist(newstate), parms = newparms, func = WHITEBUTTON, times = outtimes)

   if (input$defaultRun) {
      plot (out, defaultprog, mfrow=c(4,3), which=c("BACTERIA", "BACTERIA.13C", "FUNGI", "FUNGI.13C", "A.BISPORUS", "A.BISPORUS.13C", "SUGARS", "SUGARS.13C", "COMPOST", "CO2", "CO2.13C"), lwd = 2, las = 1, lty = 1, cex.main = 1.5, cex.axis = 1.25, cex.lab = 1.25)
      legend("bottomright", legend = c("new", "default"), cex = 1.5, col = 1:2, lty = 1)
    } else  
      plot (out, lwd = 2, las = 1, lty = 1, cex.main = 1.5, cex.axis = 1.25, cex.lab = 1.25) 
   })                          
}     

#-----------------------------------------------run as interactive user interface-----------------------------------------------

shinyApp(ui, server)

##---------------------------------------------end block interactive user interface---------------------------------------------
```