---
<<<<<<< HEAD
title: "Working model without nematodes, + literature parameters, + interactive UI, - calibrated parameters, + literature statevar."
author: "Lucy"
date: "11/04/2022"
=======
title: "Working model without nematodes, + literature parameters, + interactive UI, - calibrated parameters, - literature state variables."
author: "Lucy"
date: "27/03/2022"
>>>>>>> f4206c48fa5e4b5cddc8f479e1b717ecd7376adb
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model setup

Loading the libraries necessary, deSolve for the run of the model and shiny for the interactive UI.

```{r}
require(shiny)
require(deSolve)
```

Definition of initial state variables, initial parameters for the default model run.

```{r}
#----------------------------------------initial state variables----------------------------------------
<<<<<<< HEAD
C.INI <- 15 #300/180*6, 12-14
state <- c(BACTERIA = 0.022*C.INI, FUNGI = 0.0177*C.INI, A.BISPORUS = 0.01*C.INI, SUGARS = 0.26*C.INI, COMPOST = C.INI, CO2 = 0)
=======
C.INI <- 300/180*6
state <- c(BACTERIA = 0.10*C.INI, FUNGI = 0.05*C.INI, A.BISPORUS = 0.001*C.INI, SUGARS = 10*0.01*C.INI, COMPOST = C.INI, CO2 = 0)
>>>>>>> f4206c48fa5e4b5cddc8f479e1b717ecd7376adb

#------------------------parameters that comprise rate laws of processes go here------------------------
parms <- list(
  k1.deg.bac          = 0.05,             #0.05 - 0.15, literature value
  k2.deg.fun          = 0.05,             #0.05 - 0.15, literature value
<<<<<<< HEAD
  k3.deg.bisp         = 0.01,             #0.075 - 0.13, literature value
  k4.bac.uptake       = 0.4,              #0.04 - 0.1, literature value
  k5.fun.uptake       = 0.4,              #0.01 - 0.05, literature value
  k6.bisp.uptake      = 0.2,              #needs calibration, actually unknown, placeholder value used
  k7.fun.killing.bac  = 0.2,              #needs calibration, actually unknown, placeholder value used
  k8.bisp.killing.bac = 0.1,              #needs calibration, actually unknown, placeholder value used
  k9.bisp.killing.fun = 0.5,              #needs calibration, actually unknown, placeholder value used
  k10.bac.mort        = 0.24,             #0.24 - 0.72, literature value
  k11.fun.mort        = 0.1,              #0.01 - 0.1, literature value
  k12.bisp.mort       = 0,                #literature value
  kBISPORUS           = 0.5,              #needs calibration, actually unknown, placeholder value used
  MAX.A.BISPORUS      = 1,                #needs calibration, actually unknown, placeholder value used
  kSUGARS.bac         = 0.4,              #needs calibration, actually unknown, placeholder value used
  kSUGARS.fun         = 0.4,              #needs calibration, actually unknown, placeholder value used
  kSUGARS.bisp        = 0.2,              #needs calibration, actually unknown, placeholder value used
  Ef.fun              = 0.3,              #literature value
  Ef.bac              = 0.3,              #literature value
  Ef.bisp             = 0.4               #literature value
=======
  k3.deg.bisp         = 0.075,            #0.075 - 0.13, literature value
  k4.bac.uptake       = 0.04,             #0.04 - 0.1, literature value
  k5.fun.uptake       = 0.01,             #0.01 - 0.05, literature value
  k6.bisp.uptake      = 1,                #needs calibration, actually unknown, placeholder value used
  k7.fun.killing.bac  = 0.1,              #needs calibration, actually unknown, placeholder value used
  k8.bisp.killing.bac = 0.1,              #needs calibration, actually unknown, placeholder value used
  k9.bisp.killing.fun = 0.1,              #needs calibration, actually unknown, placeholder value used
  k10.bac.mort        = 0.24,             #0.24 - 0.72, literature value
  k11.fun.mort        = 0.01,             #0.01 - 0.1, literature value
  k12.bisp.mort       = 0,                #literature value
  kBISPORUS           = 0.1,              #needs calibration, actually unknown, placeholder value used
  MAX.A.BISPORUS      = 1.5,              #needs calibration, actually unknown, placeholder value used
  kSUGARS.bac         = 0.1,              #needs calibration, actually unknown, placeholder value used
  kSUGARS.fun         = 0.1,              #needs calibration, actually unknown, placeholder value used
  kSUGARS.bisp        = 0.1,              #needs calibration, actually unknown, placeholder value used
  Ef.fun              = 0.3,              #literature value
  Ef.bac              = 0.3,              #literature value
  Ef.bisp             = 0.3               #literature value
>>>>>>> f4206c48fa5e4b5cddc8f479e1b717ecd7376adb
)
```

Definition of the model function and the output of the model.

```{r}
WHITEBUTTON <- function(t, state, parameters) { 
  with(as.list(c(state, parameters)),{
 
#-------------Rate expressions - all in units of [mmolC/m3/day] or [molC/kg compost] go here-------------
      R1.deg.bac          <- k1.deg.bac * BACTERIA
      R2.deg.fun          <- k2.deg.fun * FUNGI
      R3.deg.bisp         <- k3.deg.bisp * A.BISPORUS
      R4.bac.uptake       <- k4.bac.uptake * SUGARS/(SUGARS+kSUGARS.bac) * BACTERIA * kBISPORUS/(A.BISPORUS+kBISPORUS)
      R5.fun.uptake       <- k5.fun.uptake * SUGARS/(SUGARS+kSUGARS.fun) * FUNGI * kBISPORUS/(A.BISPORUS+kBISPORUS)
      R6.bisp.uptake      <- k6.bisp.uptake * SUGARS/(SUGARS+kSUGARS.bisp) * A.BISPORUS * (1-A.BISPORUS/MAX.A.BISPORUS)
      R7.fun.killing.bac  <- k7.fun.killing.bac * BACTERIA * FUNGI
      R8.bisp.killing.bac <- k8.bisp.killing.bac * BACTERIA * A.BISPORUS
      R9.bisp.killing.fun <- k9.bisp.killing.fun * FUNGI * A.BISPORUS
      R10.bac.mort        <- k10.bac.mort * BACTERIA
      R11.fun.mort        <- k11.fun.mort * FUNGI
      R12.bisp.mort       <- k12.bisp.mort * A.BISPORUS
      
      #labelled example - x13C.bac <- BACTERIA13/BACTERIA
      
#---------------------------Mass balances [molC/m3/day or molC/kg/day] go here---------------------------
      dSUGARS     <- R1.deg.bac + 
                     R2.deg.fun + 
                     R3.deg.bisp -
                     R4.bac.uptake - 
                     R5.fun.uptake - 
                     R6.bisp.uptake + 
                     R7.fun.killing.bac + 
                     R8.bisp.killing.bac + 
                     R9.bisp.killing.fun +
                     R10.bac.mort + 
                     R11.fun.mort + 
                     R12.bisp.mort
      
      #labelled example - dSUGARS13 <- R20.bac.mort*x13C.BACTERIA + ... - R11.bac.uptake*x13C.SUGARS
      
      dBACTERIA   <- Ef.bac*R4.bac.uptake - 
                     R7.fun.killing.bac -
                     R8.bisp.killing.bac - 
                     R10.bac.mort 
                     
      dFUNGI      <- Ef.fun*R5.fun.uptake - 
                     R9.bisp.killing.fun - 
                     R11.fun.mort
      
      dA.BISPORUS <- Ef.bisp*R6.bisp.uptake - 
                     R12.bisp.mort
      
      #labelled example - dA.BISPORUS <- Ef.bisp*R13.bisp.uptake*x13C.SUGARS + ...
        
      dCO2        <- (1-Ef.bac)*R4.bac.uptake + 
                     (1-Ef.fun)*R5.fun.uptake + 
                     (1-Ef.bisp)*R6.bisp.uptake
      
      dCOMPOST    <- -R1.deg.bac - R2.deg.fun - R3.deg.bisp
      
      TOTAL_C     <- COMPOST + BACTERIA + SUGARS + FUNGI + A.BISPORUS + CO2
      
#----------------------------------------------Model output----------------------------------------------
      return (list(c(dBACTERIA, dFUNGI, dA.BISPORUS, dSUGARS, dCOMPOST, dCO2),
                   TOTAL_C         = TOTAL_C, 
                   R4.bac.uptake   = R4.bac.uptake,
                   R5.fun.uptake   = R5.fun.uptake,
                   R6.bisp.uptake  = R6.bisp.uptake
                   )
              )
      })
}
```

Model is ran and the results are plotted separately - argument which defines the output to model change of individually selected parameters that wish to be tracked through the duration of the dynamic model.

```{r}
outtimes     <- seq(from = 0, to = 40, length.out = 1000)
model_output <- ode(y = state, parms = parms, func = WHITEBUTTON, times = outtimes)
```

```{r, fig.height=4, fig.width=8}
plot(model_output, mfrow=c(1,3), which=c("BACTERIA", "FUNGI", "A.BISPORUS"))
```

```{r, fig.height=4, fig.width=8}
plot(model_output, mfrow=c(1,3), which=c("SUGARS", "COMPOST", "CO2"))
```

```{r, fig.height=4, fig.width=8}
plot(model_output, mfrow=c(1,3), which=c("TOTAL_C", "R4.bac.uptake", "R5.fun.uptake"))
```

```{r, fig.height=4, fig.width=8}
plot(model_output, mfrow=c(1,3), which=c("R6.bisp.uptake"))
```

```{r}
#---------------------------------------------------------------------------------------------------------
state       <- state
times       <- outtimes
defaultprog <- ode(y = state, times = outtimes, func = WHITEBUTTON, parms = parms)
#---------------------------------------------------------------------------------------------------------
UI.Si <- shinyUI(pageWithSidebar(
  headerPanel("Compost Food Web Interactive Simulation"),

  sidebarPanel(
   sliderInput(inputId="k1.deg.bac",   label = "degradation into sugars by bacteria, /d",   
     min = 0.01, max = 1.0, step = 0.01, value = parms$k1.deg.bac),
   sliderInput(inputId="k2.deg.fun",   label = "degradation into sugars by fungi, /d",        
     min = 0.01, max = 1.0, step = 0.01, value = parms$k2.deg.fun),
   sliderInput(inputId="k3.deg.bisp",  label = "degradation into sugars by A. bisporus, /d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k3.deg.bisp),
   sliderInput(inputId="k4.bac.uptake", label = "maximum carb uptake rate by bacteria, /d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k4.bac.uptake),
   sliderInput(inputId="k5.fun.uptake", label = "maximum carb uptake rate by fungi, /d",   
     min = 0.01, max = 1.0, step = 0.01, value = parms$k5.fun.uptake),
   sliderInput(inputId="k6.bisp.uptake", label = "maximum carb uptake rate by A. bisporus, /d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k6.bisp.uptake),
   sliderInput(inputId="k7.fun.killing.bac", label = "maximum predation ct fungi > bacteria, /d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k7.fun.killing.bac),
   sliderInput(inputId="k8.bisp.killing.bac", label = "maximum predation ct A. bisporus > bacteria, /d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k8.bisp.killing.bac),
   sliderInput(inputId="k9.bisp.killing.fun", label = "maximum predation ct A. bisporus > fungi, /d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k9.bisp.killing.fun),
   sliderInput(inputId="k10.bac.mort", label = "bacterial mct, (mmol C per cubic meter)/d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k10.bac.mort),
   sliderInput(inputId="k11.fun.mort", label = "fungal mct, (mmol C per cubic meter)/d",   
     min = 0.01, max = 1.0, step = 0.01, value = parms$k11.fun.mort),
   sliderInput(inputId="k12.bisp.mort", label = "A. bisporus mct, (mmol C per cubic meter)/d", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$k12.bisp.mort),
   sliderInput(inputId="kBISPORUS", label = "Monod ct A. bisporus, mmol C per cubic meter",   
     min = 0.01, max = 1.0, step = 0.01, value = parms$kBISPORUS),
   sliderInput(inputId="MAX.A.BISPORUS", label = "maxunlim A. bisporus growth, mmol C per cubic meter", 
     min = 0.01, max = 5.0, step = 0.1, value = parms$MAX.A.BISPORUS),
   sliderInput(inputId="kSUGARS.bac", label = "bacterial growth limit, mmol C per cubic meter",
     min = 0.01, max = 2.0, step = 0.01, value = parms$kSUGARS.bac),
   sliderInput(inputId="kSUGARS.fun", label = "fungal growth limit, mmol C per cubic meter", 
     min = 0.01, max = 2.0, step = 0.01, value = parms$kSUGARS.fun),
   sliderInput(inputId="kSUGARS.bisp", label = "A. bisporus growth limit, mmol C per cubic meter", 
     min = 0.01, max = 2.0, step = 0.01, value = parms$kSUGARS.bisp),
   sliderInput(inputId="Ef.bac", label = "%C assimilated by bacteria due to eating, (-)", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$Ef.bac),
   sliderInput(inputId="Ef.fun", label = "%C assimilated by fungi due to eating, (-)", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$Ef.fun),
   sliderInput(inputId="Ef.bisp", label = "%C assimilated by A. bisporus due to eating, (-)", 
     min = 0.01, max = 1.0, step = 0.01, value = parms$Ef.bisp),
   
   actionButton (inputId="resetButton", label="Reset Parameters"),
    
   checkboxInput(inputId="defaultRun",  label=strong("Add default run?"), value=TRUE),
   br()
  ),

  mainPanel(
      plotOutput("PlotSi"))
))

#-------------------------------------------------------------------------------------------------------
Server.Si <- shinyServer(function(input, output, session) {

#----------------------------------------Reset button definition----------------------------------------
  observeEvent(input$resetButton, {
    updateNumericInput(session, "k1.deg.bac",             value = parms$k1.deg.bac)
    updateNumericInput(session, "k2.deg.fun",             value = parms$k2.deg.fun)
    updateNumericInput(session, "k3.deg.bisp",            value = parms$k3.deg.bisp)
    updateNumericInput(session, "k4.bac.uptake",          value = parms$k4.bac.uptake)
    updateNumericInput(session, "k5.fun.uptake",          value = parms$k5.fun.uptake)
    updateNumericInput(session, "k6.bisp.uptake",         value = parms$k6.bisp.uptake)
    updateNumericInput(session, "k7.fun.killing.bac",     value = parms$k7.fun.killing.bac)
    updateNumericInput(session, "k8.bisp.killing.bac",    value = parms$k8.bisp.killing.bac)
    updateNumericInput(session, "k9.bisp.killing.fun",    value = parms$k9.bisp.killing.fun)
    updateNumericInput(session, "k10.bac.mort",           value = parms$k10.bac.mort)
    updateNumericInput(session, "k11.fun.mort",           value = parms$k11.fun.mort)
    updateNumericInput(session, "k12.bisp.mort",          value = parms$k12.bisp.mort)
    updateNumericInput(session, "kBISPORUS",              value = parms$kBISPORUS)
    updateNumericInput(session, "MAX.A.BISPORUS",         value = parms$MAX.A.BISPORUS)
    updateNumericInput(session, "kSUGARS.bac",            value = parms$kSUGARS.bac)
    updateNumericInput(session, "kSUGARS.fun",            value = parms$kSUGARS.fun)
    updateNumericInput(session, "kSUGARS.bisp",           value = parms$kSUGARS.bisp)
    updateNumericInput(session, "Ef.bac",                 value = parms$Ef.bac)
    updateNumericInput(session, "Ef.fun",                 value = parms$Ef.fun)
    updateNumericInput(session, "Ef.bisp",                value = parms$Ef.bisp)
  })

#-------------------Acquires model parameters which are defined in the user interface-------------------
  getparms <- reactive( {
    newparms                        <- parms 
    newparms$k1.deg.bac             <- input$k1.deg.bac
    newparms$k2.deg.fun             <- input$k2.deg.fun   
    newparms$k3.deg.bisp            <- input$k3.deg.bisp 
    newparms$k4.bac.uptake          <- input$k4.bac.uptake  
    newparms$k5.fun.uptake          <- input$k5.fun.uptake
    newparms$k6.bisp.uptake         <- input$k6.bisp.uptake 
    newparms$k7.fun.killing.bac     <- input$k7.fun.killing.bac
    newparms$k8.bisp.killing.bac    <- input$k8.bisp.killing.bac 
    newparms$k9.bisp.killing.fun    <- input$k9.bisp.killing.fun 
    newparms$k10.bac.mort           <- input$k10.bac.mort
    newparms$k11.fun.mort           <- input$k11.fun.mort
    newparms$k12.bisp.mort          <- input$k12.bisp.mort 
    newparms$kBISPORUS              <- input$kBISPORUS
    newparms$MAX.A.BISPORUS         <- input$MAX.A.BISPORUS 
    newparms$kSUGARS.bac            <- input$kSUGARS.bac
    newparms$kSUGARS.fun            <- input$kSUGARS.fun 
    newparms$kSUGARS.bisp           <- input$kSUGARS.bisp
    newparms$Ef.bac                 <- input$Ef.bac 
    newparms$Ef.fun                 <- input$Ef.fun
    newparms$Ef.bisp                <- input$Ef.bisp 
    newparms
  })

#------------------------------------------Defines the plot tab------------------------------------------
  output$PlotSi <- renderPlot({

   newparms     <- getparms()                    
   out          <- ode(y = state, parms = newparms, func = WHITEBUTTON, times = outtimes)

   if (input$defaultRun) {
      plot (out, defaultprog, which=c("BACTERIA", "FUNGI", "A.BISPORUS", "SUGARS", "COMPOST", "CO2"), lwd = 2, las = 1, lty = 1,                        cex.main = 1.5, cex.axis = 1.25, cex.lab = 1.25)  
      legend("bottomright", legend = c("new", "default"), cex = 1.5, col = 1:2, lty = 1)
    } else  
      plot (out, lwd = 2, las = 1, lty = 1, cex.main = 1.5, cex.axis = 1.25, cex.lab = 1.25)  
   })                          
})     

#--------------------------------------------------------------------------------------------------------
shinyApp(ui = UI.Si, server = Server.Si)
#----------------------------------------hacks to display better-----------------------------------------

#in div class="form-group shiny-input-container" add
# .form-group{
#  padding-top: 10px;
# }

#in label class="control-label", or essentially the one below the above container add
# .control-label{
#  margin-bottom: 3em;
# }
#--------------------------------------------------------------------------------------------------------
<<<<<<< HEAD
```
=======
```
>>>>>>> f4206c48fa5e4b5cddc8f479e1b717ecd7376adb
